Probability

### 2016-04-25 prob.tex
prob.tex several quick fix and comments...

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases. Therefore, in a resource-aware setting, it is
more appropriate to consider probability measures that map events to a
finite set of elements computable with a fixed set of
resources~\cite{Artstein1972,PuriRalescu1983}. The simplest such set,
and the one we will consider exclusively in this paper, is the set
$\mathscr{L}_{2}=\left\{ \imposs, \poss\right\}$ together with the
operation $\vee$ where $x \vee y=\imposs$ if and only if
$x=y=\imposs$. In relation to the first definition, one can interpret
\imposs\ as the closed interval $[0,0]$, \poss\ as the open interval
$(0,1]$, and $\vee$ as the addition of intervals. The definition of a
probability measure in this case is modified as being a function
$\pmeas : \events \rightarrow \mathscr{L}_{2}$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = \poss$, and
\item for a collection $E_i$,  of pairwise disjoint events, $\pmeas(\bigcup E_i) = \bigvee \pmeas(E_i)$.
\end{itemize}

@@

\begin{definition}[Quantum Probability Space~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}]
  Given a Hilbert space $\Hilb$, a \emph{quantum probability space}
  consists of a set of events $\events$ and a probability measure
  $\mu : \events \rightarrow [0,1]$ such that:\footnote {It is
    possible to define a more general space of events consisting of
    all operators~$\mathcal{A}$ on $\Hilb$ and consider
    $\mu:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.  When
    an operator $A\in\mathcal{A}$ is Hermitian, $\mu\left(A\right)$ is
    the expectation value of $A$. We does not take this approach
    because we want to focus only on probability. }
\begin{itemize}
\item The set of events consists of the empty projection $\mathbb{0}$
  together with projection operators $\proj{\psi}$ for each state
  $\ket{\psi}$,
\item $\mu(\mathbb{1})=1$, and 
\item for projections expressible as sums $\sum_{i}E_{i}$ of
  necessarily orthogonal events $E_i$, we have
  $\mu\left(\sum_{i}E_{i}\right)=\sum_{i}\mu\left(E_{i}\right)$.
\end{itemize}
\qed\end{definition}

@@

For a given set of events $\events$, there are many possible
probability measures $\mu : \events \rightarrow [0,1]$. The Born rule,
a postulate of quantum mechanics, states that each quantum state
$\ket{\phi}$ induces a probability measure $\mu_\phi$ as follows:
\[ 
\mu_\phi(E) = \ip{\phi}{E\phi}
\]
Conversely, Gleason's theorem states that given a probability measure
$\mu$, there exist a quantum state $\ket{phi}$ that induces such a
measure using the Born rule. The theorem is only valid in Hilbert
spaces with dimension $d \geq 3$. It is instructive to study
counterexamples in $d=2$, i.e., the case of a one-qubit
system. Consider five states $\ket{\psi_0}$ to $\ket{\psi_4}$ that
form five orthogonal bases $\{ \ket{\psi_0}, \ket{\psi_1} \}$,
$\{ \ket{\psi_1}, \ket{\psi_2} \}$,
$\{ \ket{\psi_2}, \ket{\psi_3} \}$,
$\{ \ket{\psi_3}, \ket{\psi_4} \}$, and
$\{ \ket{\psi_4}, \ket{\psi_0} \}$ and consider the probability
measure defined as follows. For all $i \in \{0,1,2,3,4\}$, we have
$\mu_X(\proj{\psi_i}) = 1/2$. For each orthogonal basis, the
probability is 1 as desired and yet it is impossible to find a single
quantum state that realizes such a probability
measure (see \url{http://tph.tuwien.ac.at/~svozil/publ/2006-gleason.pdf})

@@

>>>>

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases. \\
\yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
using base 16 directly.}\\
Therefore, in a resource-aware setting, it is
more appropriate to consider probability measures that map events to a
finite set of elements computable with a fixed set of
resources~\cite{Artstein1972,PuriRalescu1983}. The simplest such set,
and the one we will consider exclusively in this paper, is the set
$\mathscr{L}_{2}=\left\{ \imposs, \poss\right\}$ together with the
operation $\vee$ where $x \vee y=\imposs$ if and only if
$x=y=\imposs$. In relation to the first definition, one can interpret
\imposs\ as the closed interval $[0,0]$, \poss\ as the half-open interval
$(0,1]$, and $\vee$ as the addition of intervals. The definition of a
probability measure in this case is modified as being a function
$\pmeas : \events \rightarrow \mathscr{L}_{2}$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = \poss$, and
\item for a collection $E_i$,  of pairwise disjoint events, $\pmeas(\bigcup E_i) = \bigvee \pmeas(E_i)$.
\end{itemize}

@@

\begin{definition}[Quantum Probability Space~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}]
  Given a Hilbert space $\Hilb$, a \emph{quantum probability space}
  consists of a set of events $\events$ and a probability measure
  $\mu : \events \rightarrow [0,1]$ such that:\footnote {It is
    possible to define a more general space of events consisting of
    all operators~$\mathcal{A}$ on $\Hilb$ and consider
    $\mu:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.  When
    an operator $A\in\mathcal{A}$ is Hermitian, $\mu\left(A\right)$ is
    the expectation value of $A$. We does not take this approach
    because we want to focus only on probability. }
\begin{itemize}
\item The set of events consists of the empty projection $\mathbb{0}$
  together with projection operators $\proj{\psi}$ for each state
  $\ket{\psi}$,\\
  \yutsung{and the sum of the orthogonal events recursively.}
\item $\mu(\mathbb{1})=1$, and 
\item for projections expressible as sums $\sum_{i}E_{i}$ of
  necessarily orthogonal events $E_i$, we have
  $\mu\left(\sum_{i}E_{i}\right)=\sum_{i}\mu\left(E_{i}\right)$.
\end{itemize}
\qed\end{definition}

@@

For a given set of events $\events$, there are many possible
probability measures $\mu : \events \rightarrow [0,1]$. The Born rule,
a postulate of quantum mechanics, states that each quantum state
$\ket{\phi}$ induces a probability measure $\mu_\phi$ as follows:
\[ 
\mu_\phi(E) = \ip{\phi}{E\phi}
\]
Conversely, Gleason's theorem states that given a probability measure
$\mu$, there exist a quantum state $\ket{\phi}$ that induces such a
measure using the Born rule. The theorem is only valid in Hilbert
spaces with dimension $d \geq 3$. It is instructive to study
counterexamples in $d=2$, i.e., the case of a one-qubit
system. Consider five states $\ket{\psi_0}$ to $\ket{\psi_4}$ that
form five orthogonal bases $\{ \ket{\psi_0}, \ket{\psi_1} \}$,
$\{ \ket{\psi_1}, \ket{\psi_2} \}$,
$\{ \ket{\psi_2}, \ket{\psi_3} \}$,
$\{ \ket{\psi_3}, \ket{\psi_4} \}$, and
$\{ \ket{\psi_4}, \ket{\psi_0} \}$ and consider the probability
measure defined as follows. For all $i \in \{0,1,2,3,4\}$, we have
$\mu_X(\proj{\psi_i}) = 1/2$. For each orthogonal basis, the
probability is 1 as desired and yet it is impossible to find a single
quantum state that realizes such a probability
measure (see \url{http://tph.tuwien.ac.at/~svozil/publ/2006-gleason.pdf})

@@

\yutsung{Isn't $\proj{0}$ and $\proj{\ps}$ an example in the example of
One-qubit quantum probability space.}
